---
title: "DDP-Shiny Weather App"
author: "Maciej Nasinski"
date: "November 12, 2016"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##Introduction

```{r,echo=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(forecast)
library(xts)
```

During the Coursera Data Science course (Developing Data Products) it was built a Shiny application.
This application shows a spectrum of weather conditions across the world.
Another objective was to give a possibility to predict a future weather conditions.
Database was built using [WunderGround](https://www.wunderground.com). 
It is important to state that the application is made for educational purpose and the data should not be use for commercial purpose.
There was taken into account a continent location and a Hemisphere during selecting Cities for Database.
From a wide palette of weather indicators there was chosen only 4 of them:

- Mean Temperature Celsius
- Mean Humidity
- Mean Pressure h-Pa
- Mean Wind Speed Km-h

It should be surprising that there is not indicator for raining. 
The decision was taken because of high unpredictable of raining and lack of records for this indicator.

##Prediction Functionality

The Main functionality of the app is to assess predictions of 4 most important indicators. 
A prediction is calculated by functions from `r 'forecast'` package. 
In most cases a significant seasonality gives chance for building valuable prediction.   
Example for MeanHumidity Berlin:

```{r,echo=FALSE}
temp_all = read.csv("WeatherDatasFinal.csv")
temp_all$GMT = as.Date(temp_all$GMT)
temps = temp_all %>% group_by(GMT,year,month,week) 
tempsg = temps %>% gather(key,value,-GMT,-year,-month,-week) 
datas_pred = temps %>% group_by(year,month)%>% select(-week) %>% summarise_all(funs(mean(.,na.rm=TRUE))) 
```

```{r,echo=TRUE,fig.width=9,fig.height=4}
Holt = HoltWinters(na.approx(ts(unlist(datas_pred[,'MeanHumidity_Berlin']),frequency = 12,start=c(1996,1),end=c(2016,11))))
autoplot(forecast(Holt,h=12))
```

##Weather Conditions during Year - Plots

Many people are interested in Weather conditions across the world. 
The main aim is to show the difference in a clear manner. 
Thus it was used a one frame plot of average values of last 20 records - during the year (week interval). 
There is produced a static plot for a certain indicator and city.   
Example for MeanHumidity Berlin:

```{r cars, echo = FALSE}
temp_all = read.csv("WeatherDatasFinal.csv")
temp_all$GMT = as.Date(temp_all$GMT)
tempsg = temp_all %>% 
  group_by(GMT,year,month,week)  %>%
  gather(key,value,-GMT,-year,-month,-week) 
tempsg = tempsg%>%
  as.data.frame() %>%
  slice(grep("MeanHumidity", tempsg$key))
```

```{r,echo=TRUE,fig.width=8,fig.height=3,warning=FALSE}
data_temporary = tempsg %>% as.data.frame() %>% slice(grep("MeanHumidity_Berlin", tempsg$key)) %>%
  na.omit()
ggplot(tempsg,aes(x=week,y=value,color=key))+ geom_smooth(size=1)+
  geom_smooth(data=data_temporary,aes(x=week,y=value,color=key),size=4)+
  theme(legend.position="top")
```

##Further possibilities to develop app.

Current app is sufficient but far from being final product. 
The main aim will be to build real-time downloading of a certain city. 
Process of downloading data spread waiting time but new functionality probably worth it.

##End

Thank you for your time.

[Application](https://polkas.shinyapps.io/ShinyAppWeather/)

[Source Code](https://github.com/Polkas/ShinyWeatherApp)


